---
title: "Biostatistics 203B Final Project"
author: Jaimie and Stephanie
format: 
  html:
    theme: cosmo
    embed-resources: true
    number-sections: false
    toc: true
    toc-depth: 4
    toc-location: left
    code-fold: false
---

Display machine information:

```{r}
sessionInfo() 
```

Display my machine information:
```{r}
memuse::Sys.meminfo()
```

Load database libraries and the tidyverse frontend:

```{r}
library(bigrquery)
library(dbplyr)
library(DBI)
library(gt)
library(gtsummary)
library(tidyverse)
```

Connect to BigQuery

Authenticate with BigQuery using the service account token. Please place the service account token (shared via BruinLearn) in the working directory (same folder as your qmd file). Do **not** ever add this token to your Git repository. If you do so, you will lose 50 points.

```{r}
# path to the service account token 
satoken <- "biostat-203b-2026-winter-92fefbfab477.json"
# BigQuery authentication using service account
bq_auth(path = satoken)
```
Connect to BigQuery database `mimiciv_3_1` in GCP (Google Cloud Platform), using the project billing account `biostat-203b-2025-winter`.
```{r}
# connect to the BigQuery database `biostat-203b-2025-mimiciv_3_1`
con_bq <- dbConnect(
    bigrquery::bigquery(),
    project = "biostat-203b-2025-winter",
    dataset = "mimiciv_3_1",
    billing = "biostat-203b-2025-winter"
)
con_bq
```
List all tables in the `mimiciv_3_1` database.
```{r}
dbListTables(con_bq)
```

Connect to necessary tables.

```{r}
# full ICU stays table
icustays_tble <- tbl(con_bq, "icustays") 
# admissions data
admissions_tble <- tbl(con_bq, "admissions")
# patients data
patients_tble <- tbl(con_bq, "patients") 
# diagnoses_icd data
diagnoses_tble <- tbl(con_bq, "diagnoses_icd")

chartevents <- tbl(con_bq, "chartevents")
```

Identify hypergylcemic crisis diagnoses using ICD-9 codes.

```{r}
# Step 2: Identify hyperglycemic crisis diagnoses
hg_crisis_hadm <- diagnoses_tble |>
  filter(
    # DKA - ICD10
    icd_code %like% "E101%" |
    icd_code %like% "E111%" |
    icd_code %like% "E131%" |
    icd_code %like% "E141%" |
     # HHS ICD10
    icd_code %like% "E100%" |
    icd_code %like% "E110%" |
    icd_code %like% "E130%" |
    icd_code %like% "E140%" |
      
    # DKA - ICD9
    icd_code %like% "2501%" |
    # HHS - ICD9
    icd_code %like% "2502%"
  ) |>
  select(subject_id, hadm_id) |>
  distinct() 

#Deleted: Collect, not to collect too early in case fo merging
```

**New**: Adding outcome variable and episodes of high glucose measurements during ICU stay: 

```{r}
# Find the first day maximum glucose episode
glucose_ids <- c(220621,220615)

first_day_glucose <- chartevents |>
  filter(itemid %in% glucose_ids) |>
  select(stay_id, charttime, valuenum) |>
  inner_join(
    icustays_tble |> select(stay_id, intime),
    by = "stay_id"
  ) |>
  filter(
    charttime >= intime,
    charttime <= intime + dbplyr::sql("INTERVAL 24 HOUR")
  ) |>
  group_by(stay_id) |>
  summarise(max_glucose = max(valuenum, na.rm = TRUE)) |>
  mutate(
    hbg_group = case_when(
      max_glucose < 140 ~ "No Hyperglycemia",
      max_glucose < 200 ~ "Mild Hyperglycemia",
      TRUE ~ "Severe Hyperglycemia"
    )
  )

# Add/ Create the outcome Variable
icu_readmit <- icustays_tble |>
  group_by(subject_id) |>
  window_order(intime) |>
  mutate(
    next_icu_intime = lead(intime)
  ) |>
  ungroup() |>
  mutate(
    days_to_next_icu = dbplyr::sql("TIMESTAMP_DIFF(next_icu_intime, outtime, DAY)")
  ) |>
  mutate(
    readmit_30d = case_when(
      days_to_next_icu <= 30 ~ 1,
      TRUE ~ 0
    )
  ) |>
  select(stay_id, readmit_30d)
```



Build cohort with our established inclusion and exclusion criteria.

```{r}
# hyperglycemic_cohort <- icustays_tble |>
#   group_by(subject_id) |>
#   slice_min(intime, n = 1) |>
#   ungroup() |>
#   left_join(admissions_tble, by = c("subject_id", "hadm_id")) |>
#   left_join(patients_tble, by = "subject_id") |>
#   filter(hospital_expire_flag == 0) |>
#   filter(!is.na(insurance)) |>
#   collect() |>
#   semi_join(hg_crisis_hadm, by = c("subject_id", "hadm_id")) |>
#   mutate(
#     age_intime = anchor_age + year(intime) - anchor_year
#   ) |>
#   filter(age_intime >= 18)
# Build cohort
hyperglycemic_cohort <- icustays_tble |>
  
  # keep ICU stays linked to hyperglycemic admissions
  semi_join(hg_crisis_hadm, by = c("subject_id", "hadm_id")) |>
  
  # choose first ICU stay during that admission
  group_by(subject_id, hadm_id) |>
  slice_min(intime, n = 1, with_ties = FALSE) |>
  ungroup() |>
  
  # add admission info
  left_join(admissions_tble, by = c("subject_id", "hadm_id")) |>
  
  # add demographics
  left_join(patients_tble, by = "subject_id") |>
  
  # Add max glucose first measurent
  left_join(first_day_glucose, by = "stay_id") |>
  
  # Add 30 day outcome variable:
  left_join(icu_readmit, by = "stay_id") |>
  
  # compute age
  mutate(
    age_intime = anchor_age + year(intime) - anchor_year
  ) |>
  
  # apply exclusions
  filter(
    age_intime >= 18,
    hospital_expire_flag == 0,
    !is.na(insurance)
  ) 
```


 Collect and work with the data: 
```{r}
hyperglycemic_cohort <- hyperglycemic_cohort %>% collect()
```




```{r}
nrow(hyperglycemic_cohort)          # ICU stays
n_distinct(hyperglycemic_cohort$subject_id)  # patients
table(hyperglycemic_cohort$insurance)
summary(hyperglycemic_cohort$age_intime)
```

```{r}
if (exists("hyperglycemic_cohort_2")) {
  hyperglycemic_cohort |>
    write_rds("hyperglycemic_cohort_2.rds", compress = "gz")
}
```



