---
title: "Biostatistics 203B Final Project"
author: Name
format: 
  html:
    theme: cosmo
    embed-resources: true
    number-sections: false
    toc: true
    toc-depth: 4
    toc-location: left
    code-fold: false
---

(2) Run this code AFTER running final_dataset_creation.qmd!


```{r}
# hyperglycemic_cohort <- readRDS("hyperglycemic_cohort_2.rds")
```

```{r}
my_data <-
  hyperglycemic_cohort %>%
  
  # lump infrequent levels
  mutate(
    first_careunit      = fct_lump_n(first_careunit, n = 4), # n=5, recvommended, did not match the numeric summary on the homework
    last_careunit       = fct_lump_n(last_careunit, n = 4),
    admission_type      = fct_lump_n(admission_type, n = 4),
    admission_location  = fct_lump_n(admission_location, n = 4),
    discharge_location  = fct_lump_n(discharge_location, n = 4)
  ) %>%
  # collapse race categories
  mutate(
    race = case_when(
      str_detect(race, "WHITE")    ~ "WHITE",
      str_detect(race, "BLACK")    ~ "BLACK",
      str_detect(race, "HISPANIC") ~ "HISPANIC",
      str_detect(race, "ASIAN")    ~ "ASIAN",
      TRUE                         ~ "Other"
    ) |>
    factor(levels = c("WHITE","BLACK","HISPANIC","ASIAN","Other"))
  ) %>%
  # create long LOS indicator
  mutate(
    los_long = los >= 2
  )

my_data$insurance[is.na(my_data$insurance)] <- "No Insurance"
```


## Descriptive Statistics

```{r}
my_data |>
  select(subject_id_x, hbg_group) |>
  distinct() |>
  count(hbg_group) |>
  ggplot(aes(x = hbg_group, y = n)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Hyperglycemia Severity on First ICU Day",
    x = "Glucose Category",
    y = "Number of ICU Stays"
  )  +  scale_fill_brewer(palette = "Blues")

my_data |>
  count(hbg_group) |>
  ggplot(aes(x = hbg_group, y = n)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Hyperglycemia Severity on First ICU Day",
    x = "Glucose Category",
    y = "Number of ICU Stays"
  )  +  scale_fill_brewer(palette = "Blues")


```


## Descriptive Tables

```{r}
library(table1)
table1(~ los_long + anchor_age + gender + race + marital_status + insurance  + admission_type | readmit_30d, data = my_data)
```





## Logistic Regression


```{r}
model_data <- my_data %>%
  mutate(
    readmit_30d = factor(readmit_30d),
    insurance = factor(insurance),
    race = factor(race),
    discharge_location = factor(discharge_location),
    hbg_group = factor(hbg_group)
  ) %>%
  select(
    readmit_30d,
    insurance,
    age_intime,
    hbg_group,
    race,
    discharge_location, 
    los_long
  ) %>%
  na.omit()

model_data$insurance <- relevel(model_data$insurance, ref = "No Insurance")
#Standard Logistic Regression
fit_logit <- glm(
  readmit_30d ~ los_long + insurance + age_intime + hbg_group + race + discharge_location,
  data = model_data,
  family = binomial
)
summary(fit_logit)

# Report summary of adjusted Odds ratios

library(broom)

logit_results <- tidy(fit_logit, exponentiate = TRUE, conf.int = TRUE)

logit_results

```


Next to add for intrepretability let's center age: 

```{r}
model_data <- model_data %>%
  mutate(
    age_centered = scale(age_intime, center = TRUE, scale = FALSE)
  )
fit_logit_centered <- glm(
  readmit_30d ~ insurance + age_centered + hbg_group + race + discharge_location,
  data = model_data,
  family = binomial
)

knitr::kable(tidy(fit_logit_centered, exponentiate = TRUE, conf.int = TRUE) %>%mutate_if(is.numeric, round, 2) %>%
  select(term, estimate, conf.low, conf.high, p.value))
```




FItting of LASSO Logistic Regression: 

```{r}
#install.packages("glmnet")
library(glmnet)

x <- model.matrix(
  readmit_30d ~ insurance + age_centered + hbg_group + race + discharge_location,
  data = model_data
)[,-1]

y <- model_data$readmit_30d


# LASSO with cross-validation
lasso_model <- cv.glmnet(
  x,
  y,
  family = "binomial",
  alpha = 1
)

# Finding the best lambda
lasso_model$lambda.min

# Select of predictors
coef(lasso_model, s = "lambda.min")
```

Checking performation metrics

```{r}
library(pROC)

pred_prob <- predict(fit_logit, type = "response")

roc_obj <- roc(model_data$readmit_30d, pred_prob)

auc(roc_obj)

library(caret)

# Ensure your actual data and predictions are factors with the same levels
actual_f <- factor(model_data$readmit_30d)
pred_f <- factor(ifelse(pred_prob > 0.5, 1, 0))

# Generate the confusion matrix
confusionMatrix(pred_f, actual_f)
```

Table of Predictors:
```{r}
knitr::kable(logit_results %>%mutate_if(is.numeric, round, 2) %>%
  select(term, estimate, conf.low, conf.high, p.value))
```



Plot of odds ratios: 

```{r}
library(ggplot2)

logit_results %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = estimate, y = term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  theme_minimal() +
  labs(
    x = "Adjusted Odds Ratio",
    y = "",
    title = "Predictors of 30-Day ICU Readmission"
  )
```

## ADjusted predicted probability

```{r}
fit_logit <- glm(
  readmit_30d ~ insurance + age_intime + hbg_group + race + discharge_location,
  data = model_data,
  family = binomial
)

#Create the prediction dataset
library(dplyr)

pred_data <- model_data %>%
  summarise(
    age_intime = mean(age_intime),
    hbg_group = names(sort(table(hbg_group), decreasing = TRUE))[1],
    race = names(sort(table(race), decreasing = TRUE))[1],
    discharge_location = names(sort(table(discharge_location), decreasing = TRUE))[1]
  ) %>%
  slice(rep(1, length(unique(model_data$insurance)))) %>%
  mutate(
    insurance = levels(model_data$insurance)
  )

#predict readmission probabilities

pred_data$pred_prob <- predict(
  fit_logit,
  newdata = pred_data,
  type = "response"
)


#plot the adjusted probabilties

library(ggplot2)

ggplot(pred_data, aes(x = insurance, y = pred_prob)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  labs(
    x = "Insurance Type",
    y = "Predicted Probability of ICU Readmission",
    title = "Adjusted 30-Day ICU Readmission Risk by Insurance"
  )
# install.packages("ggeffects")
library(ggeffects)
pred_insurance <- ggpredict(fit_logit, terms = "insurance")

ggplot(pred_insurance, aes(x = x, y = predicted)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = .1) +
  theme_minimal() +
  labs(
    x = "Insurance Type",
    y = "Predicted Probability of Readmission",
    title = "Adjusted Probability of ICU Readmission by Insurance"
  )


```


## Improving the Logistic regression model


```{r}
model_data <- my_data %>%
  mutate(
    readmit_30d = factor(readmit_30d),
    insurance = factor(insurance),
    race = factor(race),
    discharge_location = factor(discharge_location),
    hbg_group = factor(hbg_group),
    hosp_los = as.numeric(dischtime - admittime)
  ) %>%
  select(
    readmit_30d,
    insurance,
    age_intime,
    hbg_group,
    race,
    discharge_location, 
    los_long,
    hosp_los
  ) %>%
  na.omit()

model_data <- model_data %>%
  mutate(
    age_centered = scale(age_intime, center = TRUE, scale = FALSE),
  )

fit_logit2 <- glm(
  readmit_30d ~ insurance + age_centered + race +
  discharge_location + los_long + hosp_los + hbg_group,
  data = model_data,
  family = binomial
)
summary(fit_logit2)

library(pROC)

pred_prob <- predict(fit_logit2, type = "response")

roc_obj <- roc(model_data$readmit_30d, pred_prob)

auc(roc_obj)


```




Correlation heatmap:

```{r}
library(reshape2)
heatmap_data <- model_data %>%
  mutate(
    readmit_30d = as.numeric(readmit_30d),
    insurance_num = as.numeric(factor(insurance)),
    race_num = as.numeric(factor(race)),
    discharge_num = as.numeric(factor(discharge_location)),
    hbg_num = as.numeric(factor(hbg_group))
  ) %>%
  select(
    readmit_30d,
    age_intime,
    los_long,
    hosp_los,
    insurance_num,
    race_num,
    discharge_num,
    hbg_num
  )

heatmap_data <- my_data %>%
  mutate(
    readmit_30d = as.numeric(readmit_30d),
    insurance_num = as.numeric(factor(insurance)),
    race_num = as.numeric(factor(race)),
    discharge_num = as.numeric(factor(discharge_location)),
    hbg_num = as.numeric(factor(hbg_group)),
    martial_status_num = as.numeric(factor(marital_status)),
    gender_num = as.numeric(factor(gender)),
    hosp_los = as.numeric(dischtime - admittime)
  ) %>%
  select(
    readmit_30d,
    age_intime,
    los_long,
    hosp_los,
    insurance_num,
    race_num,
    discharge_num,
    hbg_num,
    martial_status_num,
    gender_num,
    max_glucose,
    starts_with("lab"),
    starts_with("vital")
  )

corr_matrix <- cor(heatmap_data, use = "complete.obs")

corr_long <- melt(corr_matrix)

ggplot(corr_long, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0
  ) +
  theme_minimal() +
  labs(
    title = "Correlation Heatmap of Model Variables",
    x = "",
    y = "",
    fill = "Correlation"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

colnames(my_data)

```


## Feature Importance

```{r}
library(glmnet)

x <- model.matrix(
  readmit_30d ~ age_intime + hbg_group + los_long + hosp_los +
  insurance + race + discharge_location,
  data = model_data
)[,-1]

y <- as.numeric(model_data$readmit_30d)

lasso_model <- cv.glmnet(x, y, family = "binomial", alpha = 1)

coefs <- coef(lasso_model, s = "lambda.min")

coefs_df <- as.data.frame(as.matrix(coefs))
coefs_df$feature <- rownames(coefs_df)
coefs_df <- coefs_df[-1, ] # Remove intercept

# Filter non-zero coefficients and sort by absolute value for plotting
coefs_df <- subset(coefs_df, s1 != 0)
coefs_df$s1_abs <- abs(coefs_df$s1)
coefs_df <- coefs_df[order(coefs_df$s1_abs, decreasing = TRUE), ]

# Plot using ggplot2 or base R
# install.packages("ggplot2")
library(ggplot2)
ggplot(coefs_df, aes(x = reorder(feature, s1_abs), y = s1)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x = "Feature", y = "Standardized Coefficient Magnitude", title = "Feature Importance from cv.glmnet") +
  theme_minimal()
```

```{r}
fit_logit2 <- glm(
  readmit_30d ~ insurance + age_centered + race +
  discharge_location + los_long + hosp_los + hbg_group,
  data = model_data,
  family = binomial
)

model_data$pred_prob <- predict(
  fit_logit2,
  type = "response"
)

library(pROC)
library(ggplot2)

roc_obj <- roc(model_data$readmit_30d, model_data$pred_prob)

auc_val <- auc(roc_obj)
auc_val
roc_df <- data.frame(
  tpr = roc_obj$sensitivities,
  fpr = 1 - roc_obj$specificities
)

ggplot(roc_df, aes(x = fpr, y = tpr)) +
  geom_line(size = 1.2, color = "steelblue") +
  geom_abline(linetype = "dashed") +
  theme_minimal() +
  labs(
    title = paste("ROC Curve (AUC =", round(auc_val,3), ")"),
    x = "False Positive Rate",
    y = "True Positive Rate"
  )

## Plot the calibration curve: 
library(dplyr)

calibration_data <- model_data %>%
  mutate(
    risk_bin = ntile(pred_prob, 10)
  ) %>%
  group_by(risk_bin) %>%
  summarise(
    predicted = mean(pred_prob),
    observed = mean(as.numeric(readmit_30d) - 1)
  )

ggplot(calibration_data, aes(predicted, observed)) +
  geom_point(size = 3, color = "darkred") +
  geom_line(color = "darkred") +
  geom_abline(linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Calibration Plot",
    x = "Predicted Readmission Probability",
    y = "Observed Readmission Rate"
  )

model_data %>%
  mutate(risk_group = ntile(pred_prob, 5)) %>%
  group_by(risk_group) %>%
  summarise(readmit_rate = mean(as.numeric(readmit_30d)-1)) %>%
  ggplot(aes(risk_group, readmit_rate)) +
  geom_col(fill = "purple") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(
    title = "Observed Readmission Rate by Predicted Risk Quintile",
    x = "Predicted Risk Quintile",
    y = "Observed Readmission Rate"
  )

```



```{r}
# library(dplyr)
# library(dbplyr)
# library(lubridate)
# 
# icustays <- tbl(con_bq, "icustays")
# admissions <- tbl(con_bq, "admissions")
# patients <- tbl(con_bq, "patients")
# diagnoses <- tbl(con_bq, "diagnoses_icd")
# chartevents <- tbl(con_bq, "chartevents")
# 
# # Find DKA and HHS Admissions
# hg_dx <- diagnoses |>
#   filter(
#     icd_code %like% "E101%" |
#     icd_code %like% "E111%" |
#     icd_code %like% "E131%" |
#     icd_code %like% "E141%" |
#     icd_code %in% c("E1100","E1101","E1300","E1301") |
#     icd_code %like% "2501%" |
#     icd_code %like% "2502%"
#   ) |>
#   select(subject_id, hadm_id) |>
#   distinct()
# 
# 
# # Create the ICU cohort
# cohort <- icustays |>
#   semi_join(hg_dx, by = c("subject_id","hadm_id")) |>
#   left_join(admissions, by = c("subject_id","hadm_id")) |>
#   left_join(patients, by = "subject_id") |>
#   mutate(
#     age_intime = anchor_age + year(intime) - anchor_year
#   ) |>
#   filter(
#     age_intime >= 18,
#     hospital_expire_flag == 0,
#     !is.na(insurance)
#   )
# 
# # Find the first day maximum glucose episode
# glucose_ids <- c(220621,220615)
# 
# first_day_glucose <- chartevents |>
#   filter(itemid %in% glucose_ids) |>
#   select(stay_id, charttime, valuenum) |>
#   inner_join(
#     cohort |> select(stay_id, intime),
#     by = "stay_id"
#   ) |>
#   filter(
#     charttime >= intime,
#     charttime <= intime + days(1)
#   ) |>
#   group_by(stay_id) |>
#   summarise(max_glucose = max(valuenum, na.rm = TRUE)) |>
#   mutate(
#     hbg_group = case_when(
#       max_glucose < 140 ~ "No Hyperglycemia",
#       max_glucose < 200 ~ "Mild Hyperglycemia",
#       TRUE ~ "Severe Hyperglycemia"
#     )
#   )
# 
# # Add/ Create the outcome Variable
# icu_readmit <- icustays |>
#   arrange(subject_id, intime) |>
#   group_by(subject_id) |>
#   mutate(
#     next_icu_intime = lead(intime),
#     days_to_next_icu = as.numeric(next_icu_intime - outtime)
#   ) |>
#   ungroup() |>
#   mutate(
#     readmit_30d = case_when(
#       days_to_next_icu <= 30 ~ 1,
#       TRUE ~ 0
#     )
#   ) |>
#   select(stay_id, readmit_30d)
# 
# 
# # Merge everything
# my_data  <- cohort |>
#   left_join(first_day_glucose, by = "stay_id") |>
#   left_join(icu_readmit, by = "stay_id") |>
#   collect()
```


### Sanity check for Dataset: 

```{r}
nrow(analysis_dataset)

table(analysis_dataset$hbg_group)

table(analysis_dataset$insurance)

mean(analysis_dataset$readmit_30d)
```


